---
title: "The supplementary file"
author: "Yiliu Cao"
date: "2023-04-06"
output: pdf_document
---

In this file, I will show the generated data in the experiment.
```{r}
# Set up the factorial design with complete blocks
# The default number of block is 2 and the number of factor is 3
n <- 50
x1 <- rep(c(-1, 1), 2^3*n)
x2 <- rep(c(-1, -1, 1, 1), 2^3*n/2)
x3 <- rep(c(rep(-1, 4),rep(1,4)), 2^3*n/4)

# Distinguish the blocks for each observation
blk <- (c(rep(1, 2^3*n), rep(2, 2^3*n)))
run <- c(1:(2^3*n*2))
data <- data.frame(Run = run,
           M = x1,
           O = x2,
           P = x3,
           Block = blk)
```

Generate the outcome:
```{r}
# Set up the outcome
set.seed(305)
# Different factor has different effects on the outcome
rating <- c()
for (i in 1:nrow(data)){
  basic <- 50
  if (data[i,2] == 1){
    basic <- basic + rnorm(1, 15, 5)
  }
  if (data[i,3] == 1){
    basic <- basic + rnorm(1, 15, 10)
  }
  if (data[i,4] == 1){
    basic <- basic + rnorm(1, 10, 5)
  }
  if (basic > 100){
    basic <- 100
  }
  else if (basic < 0){
    basic <- 0
  }
  rating[i] <- round(basic, 0)
}
data$Rating <- rating
# You can choose to export the data set
# write.csv(data, "data_FP.csv", row.names = FALSE)
```

In my codes, you can find the I set the first half observations to be the block 1 and the second half to be block 2. However, the number of observations is huge, I will only present 40 observations for each block (Runs 1 to 40 and Runs 401 to 440).\
For the block 1"
```{r}
print(data[c(1:40), ])
```

For the block 2:
```{r}
print(data[c(401:440), ])
```

If you want the full data, you can run the following code:
```{r}
# write.csv(data, "data_FP.csv", row.names = FALSE)
```
Then you can have a data named "data_FP.csv" locally.
